---
title: "Run lengths"
author: "Nic McPhee"
date: "2/3/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Run lengths

This compares the number of generations in runs using both the SMAC and "standard"
parameters on the replace-space-with-newline (RSWN) problem, both across all
runs and when limited to successful runs.

```{r}
run_lengths <- read_delim("run_lengths.txt", " ", escape_double = FALSE, trim_ws = TRUE)

successful_runs = subset(run_lengths, Success==TRUE)
```

## Make boxplots

Boxplot of run length across all runs, successful or not:

```{r echo=FALSE}
ggplot(run_lengths, aes(x=Treatment, y=Num.generations)) + geom_boxplot() + theme_bw()
```

Boxplot of run lengths limited just to successful runs:

```{r echo=FALSE}
ggplot(successful_runs, aes(x=Treatment, y=Num.generations)) + geom_boxplot() + theme_bw()
```

## Wilcoxon tests

Testing for significant difference in run length across all runs, regardless 
of whether they succeeded:

```{r}
wilcox.test(Num.generations ~ Treatment, data=run_lengths)
```
```{r}
wilcox.test(Num.generations ~ Treatment, data=successful_runs)
```

```{r}
rl_ps = data.frame(Generation=integer(), "p-value"=double())
for (g in unique(run_lengths$Num.generations)) {
  smac = nrow(subset(run_lengths, Success == TRUE & Num.generations < g & Treatment == "SMAC"))
  standard = nrow(subset(run_lengths, Success == TRUE & Num.generations < g & Treatment == "Standard"))
  new_row = data.frame(Generation=g, "p-value"=prop.test(c(smac, standard), c(110, 110))$p.value)
  rl_ps <- rbind(rl_ps, new_row)
}
rl_ps <- rl_ps[order(rl_ps$Generation),]
rl_ps$adjusted.p <- p.adjust(rl_ps$p.value)
```