\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{todonotes}

% Our added packages
\usepackage[ruled]{algorithm2e}
\usepackage{tabu}

%\usepackage{algorithm}
%\usepackage{algorithmic}


% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[GECCO '17]{the Genetic and Evolutionary Computation Conference 2017}{July 15--19, 2017}{Berlin, Germany}
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}


\begin{document}
\title{Using SMAC to (over) optimize parameters for PushGP}
\subtitle{Do we want a sub-title? I bet we can improve on this title.}

\author{Nicholas Freitag McPhee}
\orcid{0000-0002-6495-2612}
\affiliation{%
	\institution{University of Minnesota, Morris}
	% \streetaddress{600 E. 4th Street}
	\city{Morris} 
	\state{Minnesota} 
	\country{USA}
	\postcode{56267}
}
\email{mcphee@morris.umn.edu}

\author{Thomas Helmuth}
\affiliation{%
    \institution{Washington and Lee University}
    \city{Lexington} 
    \state{Virginia} 
	\country{USA}
	\postcode{24450}
}
\email{helmutht@wlu.edu}

\author{Lee Spector}
\affiliation{%
	\institution{Hampshire College}
	\city{Amherst}
	\state{Massachusetts}
	\country{USA}
	\postcode{01002}
}
\email{lspector@hampshire.edu}

\begin{abstract}
	Sequential Model-based Algorithm Configuration (SMAC) is a
	tool for optimizing algorithm parameters, using repeated runs
	of the target algorithm with different parameter values to estimate
	the relationship between parameters and performance. 
	Here we use SMAC to optimize
	a group of parameters for PushGP, a stack-based genetic programming
	system, for
	several software synthesis problems. Applying
	SMAC to one particular problem leads
	to marked improvements in the success rate and the speed with which
	a solution was found for that problem. Applying these ``tuned'' 
	parameters to four
	additional problems, however, provided no benefit on three, and
	substantially reduced performance on the last. This suggests that SMAC 
	is ``overfitting'', tuning the parameters in ways that are 
	highly problem specific, and raises doubts about the value of
	using these ``tuned'' parameters on  previously unsolved problems.
	Efforts to use SMAC to optimize PushGP parameters on other problems
	have been less successful due to a combination of long PushGP run times
	and low success rates, which make it hard for SMAC to acquire enough
	signal in a reasonable amount of time.
	
	\todo[inline]{We need to do the CCS Concepts thing.}
	\todo[inline]{We need to do the keywords.}
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

\keywords{ACM proceedings, \LaTeX, text tagging}


\maketitle

\input{SMAC-and-PushGP-body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{SMAC-and-PushGP} 

\end{document}
